---
title: "Final Project"
subtitle: "Data Science for Biologists, Fall 2021"
author: "Nikita Thadani"
output: 
   prettydoc::html_pretty:
    theme: hpstr
    highlight: pygments
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# DON'T CHANGE THE NEXT LINE, except to add NEW options if you have any, like default figure sizes
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo = TRUE, include = TRUE)

# Path to figure export directory 
path_to_figures <- file.path(here::here(), "figures")

# Load all libraries you use HERE. If you have to install something, please make sure to discuss with Spielman.
library(tidyverse)
library(pROC)
theme_set(theme_classic())
```

<br><br>

## Introduction

I obtained the Breast Cancer cell data from the Kaggle website [here](https://www.kaggle.com/uciml/breast-cancer-wisconsin-data), which also directs you to the UCI Machine Learning Repository [here](https://archive.ics.uci.edu/ml/datasets/Breast+Cancer+Wisconsin+%28Diagnostic%29). The dataset contains 569 samples (357 benign and 212 malignant) of different features of breast tumor cell nuclei by using the biopsy method of fine needle aspirate (FNA) of a breasts mass. I picked this data set mainly because I actually wanted to learn what measurements are taken into consideration when doing biopsy on breast mass, I had no idea what exactly helps determine whether it is a malignant or benign breast tumor. Quite honestly by doing the model to predict future diagnosis, makes me feel successful in a way, even though my model is not being used for such important diagnosis, I feel like I can *actually* do models to keep pushing for better healthcare systems. 

The attributes in the data set are as follows:

+ id: Identifies the sample
+ diagnosis: Tumor is 1= Malignant or 0= Benign
+ radius: Mean of distances from center to points on the perimeter 
+ texture:  Standard deviation of gray-scale values 
+ perimeter: Distance around nuclear border  
+ area : Area of nucleus 
+ smoothness: Variation in the cell’s radial lengths  
+ compactness: perimeter$^2$ / area)
+ concavity: Size of indentations in the nuclear border
+ concave points: Number of concave portions of the nuclear border
+ symmetry: Deviation of nuclei shape from ideal 
+ fractal dimension: Measurement of nuclear border irregularity or "coastline approximation"
+ The mean(`variable_mean`), standard error(`variable_se`) and "worst" or largest (mean of the three
largest values, shown as `variable_se`) of these features were computed for each image,
resulting in 30 features. All feature values are recorded with four significant digits

Since the attributes have been explained above, the `variables` in the dataset are as follow:

+ `id`
+ `diagnosis`
+ `radius_mean`
+ `texture_mean`
+ `perimeter_mean` 
+ `area_mean`
+ `smoothness_mean`
+ `compactness_mean`
+ `concavity_mean`
+ `concave points_mean`         
+ `symmetry_mean `
+ `fractal_dimension_mean`
+ `radius_se`
+ `texture_se` 
+ `perimeter_se` 
+ `area_se`
+ `smoothness_se`
+ `compactness_se`
+ `concavity_se`
+ `concave points_se`
+ `symmetry_se`
+ `fractal_dimension_se`
+ `radius_worst` 
+ `texture_worst `
+ `perimeter_worst `
+ `area_worst`
+ `smoothness_worst` 
+ `compactness_worst` 
+ `concavity_worst `
+ `concave points_worst`
+ `symmetry_worst`
+ `fractal_dimension_worst`
+ `...33 `: This variable is full of NAs, it will be later deleted




```{r read-examine-data}
# Code to read in the data:
breast_cancer <-read_csv("/cloud/project/project_proposal/thadani_dataset.csv") 

# Code to dplyr::glimpse() the data:
glimpse(breast_cancer)


#clean data

breast_cancer %>%
  select(-id) %>% 
  select(- ...33) %>%
  mutate(diagnosis  = if_else(diagnosis  == "M", 1, 0))  -> new_can
new_can


#data with full name of diagnosis

new_can %>% 
  
mutate(diagnosis  = if_else(diagnosis == 1, 
                            "Malignant", 
                            "Benign")) -> diagnosis_breast
diagnosis_breast


```

<br><br>

## Question 1: Modeling

#### Do these specific variables explain variation in/predict status of response?

### Methods

```{r model}
model_selection <- glm(diagnosis ~ radius_mean 
                       + texture_mean 
                       + perimeter_mean 
                       + area_mean 
                       + smoothness_mean 
                       + compactness_mean 
                       + concavity_mean 
                       + symmetry_mean 
                       + fractal_dimension_mean,
                       data = new_can,
                       family = "binomial")

model_formula <- as.formula(model_selection)
model_formula


model_roc <- roc(new_can$diagnosis,
                 model_selection$fitted.values)  #MODEL!
model_roc$auc  #extract auc    #0.9866


#Logic curve 

tibble(logistic_x = model_selection$linear.predictors, 
       logistic_y = model_selection$fitted.values, 
       truth = if_else(new_can$diagnosis == 1, "Malignant", "Benign")) -> logistic_model_plot_data


ggplot(logistic_model_plot_data) + 
  aes(x = logistic_x,
      y = logistic_y) + 
  geom_line() + 
  geom_point(aes(color = truth), alpha = 0.4) +
  scale_color_viridis_d(option = 'plasma') + 
  labs(x = "Linear predictors", 
       y = "Predicted probability of breast cancer", 
       color = "True disease status",
       title = "Logistic Curve") -> modelfig
modelfig

output_figure_file <- file.path(path_to_figures, 
                                "question1.png")

ggsave(output_figure_file, # output file name to save a PNG file
       modelfig,             
       width = 6,          
       height = 4)
```


### Answer

The model has an AUC of 0.9866 which indicates a strong model since it has a high AUC.
 


<br><br>

## Question 2

#### Are malignant cells more concave and have a bigger radius than benign ones?.

### Methods

```{r q2}

diagnosis_breast %>%
  ggplot() +
  aes(x= radius_mean,
      y= concavity_mean,
      color = diagnosis)+
  geom_point() +
  labs(x= "Radius",
       y = "Concavity",
       color = "Diagnosis",
       title = "Are malignant cells more concave and have a bigger radius than benign ones?")+
  geom_smooth(method = "lm",
              color = "red") +
  scale_color_viridis_d(option = 'cividis') -> q2 
q2


output_figure_file <- file.path(path_to_figures, 
                                "question2.png")

ggsave(output_figure_file, # output file name to save a PNG file
       q2,             
       width = 8,          
       height = 4)
```



### Answer

The plot shows that malignant cells are more concave and have a bigger radius than benign ones because there is a positive linear relationship between the variables. 

<br><br>

## Question 3

#### What is the relationship between the area and compactness of the cell nuclei? 

### Methods

```{r q3}
diagnosis_breast %>%
  ggplot() +
  aes(x= area_mean,
      y= compactness_mean,
      color = diagnosis)+
  geom_point() +
  labs(x= "Area",
       y= "Compactness",
       color = "Diagnosis",
       title = "What is the relationship between the area and compactness of the cell nuclei?")+
  geom_smooth(method = "lm",
              color = "yellow") +
  scale_color_brewer(palette = 'Set1')-> q3
q3


output_figure_file <- file.path(path_to_figures, 
                                "question3.png")

ggsave(output_figure_file, # output file name to save a PNG file
       q3,             
       width = 8,          
       height = 4)
```
 


### Answer

The plot shows a positive linear relationship between the area and compactness of the cell nuclei.  


<br><br>

## Question 4

#### How many malignancies are in the dataset?

### Methods

```{r q4}
hsize <- 5

diagnosis_breast %>%
  count(diagnosis) %>%
  mutate(hsize) %>%
  ggplot()+
  aes(x = hsize, 
      y= n,
      fill= diagnosis) +
  geom_col() +
  geom_text(aes(label = n),
            position = position_stack(vjust = 0.5),
            show.legend = FALSE) +
  coord_polar(theta = "y") +
  xlim(c(0.2, hsize + 0.5))+
  labs(title ="How many malignancies are in the dataset?",
       fill= "Diagnosis") +
  scale_fill_brewer(palette = 'Set3') +
  theme_void() -> q4
q4

output_figure_file <- file.path(path_to_figures, 
                                "question4.png")

ggsave(output_figure_file, # output file name to save a PNG file
       q4,             
       width = 6,          
       height = 4)

```



### Answer

There are 356 malignancies in this data set. 
